Mathematical Aspects of the Capital Market

### 2017-08-21 phd/MyThesis.tex
Revised my phd thesis.
Revised the phd thesis.

\usepackage[numbers,sort&compress]{natbib}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{appendix}
\graphicspath{
        {../papers/FX/}{../papers/Jeffrey1/}
        {../papers/Anja/}
        {../papers/DavisMikoschHeinyXie/}
}

@@

%\vspace*{1cm}
\hspace*{0.5cm}  {\bf Xiaolei Xie}\\
\hspace*{1.5cm} Department of Mathematical Sciences\\
\hspace*{1.5cm} University of Copenhagen\\
\hspace*{1.5cm} Universitetsparken 5\\
\hspace*{1.5cm} 2100 København Ø\\
\hspace*{1.5cm} Denmark\\
\hspace*{1.5cm} xie@math.ku.dk\\
\hspace*{1.5cm} xie.xiaolei@gmail.com

@@

\begin{tabular}{ll}
Supervisor: & Prof.~Thomas Mikosch\\
& University of Copenhagen\\
& \\
Co-supervisor: & Associate Prof.~Jeffrey Collamore\\
& University of Copenhagen\\
& \\

Assessment committee:  & Lecturer.~Jesper Lund Pedersen (chairman)\\
& University of Copenhagen\\
& \\

&  Prof.~Catalin St{\u{a}}ric{\u{a}}\\
& University of Neuchatel, CH \\
& \\

@@

\vspace*{5cm}
\noindent \hspace*{0.5cm} Copyright:\\
\hspace*{0.5cm} Chapters 1-4: \copyright Johannes Heiny (according to Danish legislation)\\
\hspace*{0.5cm} Chapter 2: \copyright Elsevier\\

\noindent \hspace*{0.5cm} ISBN: $978$-$87$-$7078$-$921$-$9$

@@

This thesis is about heavy-tailed time series.

\section*{Resum{\'e}}

\newpage
\section*{Acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}

I owe my gratitude to many people and especially to my supervisors Prof. Mikosch and Prof. Collamore.
\bigskip

{\em Xie Xiaolei\qquad \qquad \qquad \qquad \qquad Copenhagen, September $2017$.}

@@

\begin{enumerate}
\bibitem[P2]{davis:mikosch:heiny:xie:20151}
{\sc Davis, R.~A., Heiny, J., Mikosch, T., and Xie, X.}
\newblock Extreme value analysis for the sample autocovariance matrices of
  heavy-tailed multivariate time series.
\newblock {\em Extremes 19}, 3 (2016), 517--547. \href{http://link.springer.com/article/10.1007/s10687-016-0251-7}{\blue{[pdf]}}
\end{enumerate}

Now we explain the structure of this thesis.

{\em Xie Xiaolei \qquad \qquad \qquad \qquad \qquad Copenhagen, January $2017$.}

>>>>

\usepackage[numbers,sort&compress]{natbib}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{appendix}
\usepackage{multirow}
\graphicspath{
        {../papers/FX/}{../papers/Jeffrey1/}
        {../papers/Anja/}
        {../papers/DavisMikoschHeinyXie/}
}

@@

%\vspace*{1cm}
\hspace*{0.5cm}  {\bf Xiaolei Xie}\\
\hspace*{1.5cm} Department of Mathematical Sciences\\
\hspace*{1.5cm} University of Copenhagen\\
\hspace*{1.5cm} Universitetsparken 5\\
\hspace*{1.5cm} 2100 Copenhagen\\
\hspace*{1.5cm} Denmark\\
\hspace*{1.5cm} xie@math.ku.dk\\
\hspace*{1.5cm} xie.xiaolei@gmail.com

@@

\begin{tabular}{ll}
Supervisor: & Prof.~Thomas Mikosch\\
& University of Copenhagen\\
& \\
Co-supervisor: & Associate Prof.~Jeffrey Collamore\\
& University of Copenhagen\\
& \\
Co-supervisor: & Prof.~Rolf Poulsen\\
& University of Copenhagen\\
& \\
Assessment committee:  & Associate Prof.~Jesper Lund Pedersen (chairman)\\
& University of Copenhagen\\
& \\

&  Prof.~Catalin St{\u{a}}ric{\u{a}}\\
& University of Neuchatel\\
& \\

@@

\vspace*{5cm}
\noindent \hspace*{0.5cm} Copyright:\\
\hspace*{0.5cm} Chapters 1-2: \copyright Xiaolei Xie (according to Danish legislation)\\
\hspace*{0.5cm} Chapter 3: \copyright Bernoulli\\
\hspace*{0.5cm} Chapter 4: \copyright Elsevier\\

% \noindent \hspace*{0.5cm} ISBN: $978$-$87$-$7078$-$921$-$9$

@@

This thesis is about analysis of heavy-tailed time series. We discuss tail
properties of real-world equity return series and investigate the
possibility that a single tail index is shared by all return series of
actively traded equities in a market. Conditions for this hypothesis to
be true are identified.

We study the eigenvalues and eigenvectors of sample covariance and
sample auto-covariance matrices of multivariate heavy-tailed time series,
and particularly for time series with very high dimensions. Asymptotic
approximations of the eigenvalues and eigenvectors of such matrices
are found and expressed in terms of the parameters of the dependence
structure, among others.

Furthermore, we study an importance sampling method for estimating
rare-event probabilities of multivariate heavy-tailed time series
generated by matrix recursion. We show that the proposed algorithm is
efficient in the sense that its relative error remains bounded as the
probability of interest tends to 0. We make use of exponential
twisting of the transition kernel of an {\em Markov additive process},
and take advantage of asymptotic theories on products of positive
random matrices.

\section*{Resum{\'e}}
Denne afhandling handler om analyse af tunge tailed tidsserier. Vi
diskuterer halen Egenskaber af real-world egenkapital retur serie og
undersøge mulighed for at et enkelt haleindeks deles af alle
returserier af aktivt handlede aktier i et marked. Betingelser for
denne hypotese til Være sande er identificeret.

Vi studerer egenværdier og egenvektorer af observerede kovarians og
observerede auto-kovarians matricer af multivariate heavy-tailed
tidsserier, og især til tidsserier med meget høje dimensioner.
Asymptotisk tilnærmelser af egenværdierne og egenvektorer af
sådanne matricer findes og udtrykkes i afhængighedens parametre
struktur, blandt andre.

Vi studerer også en importance samplings metod til estimering af
sjældne begivenheder sandsynligheder for multivariate heavy-tailed
tidsserier genereret af matrix recursion. Vi viser, at den foreslåede
algoritme er effektiv i den forstand, at dens relative fejl forbliver
afgrænset som sandsynligheden af interesse har tendens til 0. Vi gør
brug af eksponentiel vridning af overgangskernen til en {\em Markov
additive process} og tage fordel af asymptotiske teorier om produkter
af positiv tilfældig matricer.
\newpage
\section*{Acknowledgments}
\addcontentsline{toc}{section}{Acknowledgments}
I owe my gratitude to many people and especially to my supervisors
Prof. Thomas Mikosch and Prof. Jeffrey Collamore. Their insight and
guidance are indispensible throughout the scientific endeavor that
culminates in this thesis. Meanwhile, I cannot be more grateful to
Prof. Mogens Steffensen -- without his financial
support from the HIPERFIT program, my PhD project would not have
been feasible.

I also feel deeply indebted to my co-authors Prof. Richard A. Davis,
Prof. Casper de Vries, Prof. Anja Jan\ss en and Dr. Johannes Heiny.
Collaboration with them is not only fruitful but enjoyable. Those moments
of relaxed discussions are to be remembered forever. Dr. Johannes Heiny
was also my office-mate. I truly enjoyed those casual conversations
with him about maths and life in general.

To the professors and PhD students in the Mathematics department of
Copenhagen University and especially to Prof. Jostein Paulsen, I owe a
big ``thank you'' too. I really enjoyed working with him as a teaching
assistant and at the same time learned a lot of fantastic maths from him
and his book manuscript.

This is also the opportunity for me to thank Dr. Dan Cheng and
Dr. Xuan Leng, whom I first met at EPFL Lausanne and later at TU Delft.
Discussions with them have been greatly helpful and inspiring to me.

Last but not least, I would like to thank the PhD committee for spending
their precious time to review and assess this thesis.
\bigskip

{\em Xiaolei Xie\qquad \qquad \qquad \qquad \qquad Copenhagen, September $2017$.}

@@

\begin{enumerate}
%% \bibitem{mikosch:deVries:xie:2016}
%% {\sc de Vries, C. and Mikosch, T. and Xie, X.}
%% \newblock Do return series have power-law tails with the same
%%   index?
%% \newblock working paper.

%% \bibitem{collamore:vidyashankar:xie:2017}
%% {\sc Collamore, Jeffrey and Vidyashankar, Anand and Xie, Xiaolei}
%% \newblock Rare event simulation for GARCH(p,q) processes.
%% \newblock working paper.

%% \bibitem{janssen:mikosch:rezapour:xie:2016}
%% {\sc Janssen, A. and Mikosch, T. and Mohsen, R. and X., Xie}
%% \newblock The eigenvalues of the sample covariance matrix of a multivariate heavy-tailed stochastic volatility model.
%% \newblock {\em Bernoulli Journal}, To appear

%% \bibitem[P4]{davis:mikosch:heiny:xie:2015}
%% {\sc Davis, R.~A., Heiny, J., Mikosch, T., and Xie, X.}
%% \newblock Extreme value analysis for the sample autocovariance matrices of heavy-tailed multivariate time series.
%% \newblock {\em Extremes 19}, 3 (2016), 517--547. \href{http://link.springer.com/article/10.1007/s10687-016-0251-7}
\bibitem[P1]{mikosch:deVries:xie:2016}
{\sc Mikosch, T., de Vries, C. and Xie, X.}
\newblock Do return series have power-law tails with the same index?
\newblock {\em Technical Report}

\bibitem[P2]{collamore:vidyashankar:xie:2017}
{\sc Collamore, J., Vidyashankar, A. and Xie, X.}
\newblock Rare event simulation for GARCH(p,q) processes.
\newblock {\em Technical Report}.

\bibitem[P3]{janssen:mikosch:rezapour:xie:2016:intro}
{\sc Janssen, A., Mikosch, T., Mohsen, R. and Xie, X.}
\newblock Limit theory for the singular values of the sample autocovariance
  matrix function of multivariate time series.
\newblock {\em Bernoulli, to appear}.

\bibitem[P4]{davis:mikosch:heiny:xie:20151}
{\sc Davis, R.~A., Heiny, J., Mikosch, T., and Xie, X.}
\newblock Extreme value analysis for the sample autocovariance matrices of
  heavy-tailed multivariate time series.
\newblock {\em Extremes 19}, 3 (2016), 517--547. \href{http://link.springer.com/article/10.1007/s10687-016-0251-7}{\blue{[pdf]}}
\end{enumerate}

%% {\em Xiaolei Xie \qquad \qquad \qquad \qquad \qquad Copenhagen, January $2017$.}
